Towards a miniKanren with fair search strategies

### 2019-06-18 Paper/paper.tex
revise

To summarize our contributions, we
\begin{itemize}
	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
depth-first search (\DFSbi{}), a new search strategy with almost-fair \disj{}.
	\item propose and implement \textbf{f}air depth-first search (\DFSf{}), 
a new search strategy with fair \disj{}.
	\item implement in a new way the standard breath-first search (\BFS) by 
\citet{seres1999algebra}, a search strategy with fair \disj{} and fair \conj{}. 
We refer to \BFSser{} as their implementation and \BFSimp{} as our new 
one. We formally prove that the two implementations are 
semantically equivalent, however, \BFSimp{} runs faster in all benchmarks and 
is shorter.
\end{itemize}

@@

In this section, we define fair \disj{}, almost-fair \disj{} and fair \conj{}. 
Before going further into fairness, we give a short review of the terms:
\emph{state}, \emph{space}, and \emph{goal}.
A \emph{state} is a collection of constraints. (Here, we restrict 
constraints to unification constraints.) Every answer corresponds to a 
state. A \emph{space} is a collection of states. And a \emph{goal} is a 
function from a state to a space.

@@

Although some strategies have a different level of fairness in \disj{}, they 
have the same behavior when there is no call to a relational definition in 
\conde{} clauses (including this case).

@@

Our solution requires a search strategy with \emph{fair} \conj{} to organize
states in buckets in spaces, where each bucket contains finite states, 
and 
to allocate resources evenly among spaces derived from states in the 
same bucket. It is up to a search strategy designer to decide by what criteria 
to 
put states in the same bucket, and how to allocate resources among spaces 
related to different buckets.

@@

\begin{table}
	% run in command line, with a memory limit of 500MB
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline 
	benchmark & size & \DFSi & \DFSbi & \DFSf & \BFSimp{} & \BFSser{} \\
		\hline
		\veryrecursiveo & 100000 & 166 & 103 & 412 &   451 &   1024 \\
	    				& 200000 & 283 & 146 & 765 &   839 &   1875 \\
						& 300000 & 429 & 346 &2085 &  1809 &   4408 \\
		\hline  
		\appendo        & 	 100 &  17 &  18 &  17 &    17 &     65 \\
		         	  	& 	 200 & 145 & 137 & 137 &   142 &    121 \\
          		      	&	 300 & 388 & 384 & 387 &   429 &    371 \\
%   		& - & - & - & - & - & - \\
		\hline 
		\reverso		& 	  10 &   3 &   4 &   3 &    18 &     36 \\
						& 	  20 &  35 &  35 &  33 &  3070 &   3695 \\
						& 	  30 & 329 & 333 & 315 & 75079 & 107531 \\
		\hline
		quine-1 		& 1 & 13 & 14 & 10 & - & - \\
						& 2 & 30 & 34 & 25 & - & - \\
						& 3 & 41 & 48 & 33 & - & - \\
		\hline
		quine-2 		& 1 & 22 & 21 & 12 & - & - \\
						& 2 & 51 & 46 & 24 & - & - \\
						& 3 & 72 & 76 & 32 & - & - \\
		\hline
		'(I love you)-1 &  99 &  10 &  37 & - &   40 &   96 \\
						& 198 &  22 &  71 & - &  374 &  652 \\
						& 297 &  24 & 292 & - & 1668 & 3759 \\
		\hline
		'(I love you)-2 &  99 & 485 & 179 & - &   44 &   94 \\
						& 198 & 808 & 638 & - &  414 &  630 \\
						& 297 &1265 & 916 & - & 1651 & 3691 \\ 
		\hline 
		% The data below is test with old benchmarks
%		quine-1 & 1 &  71 &  44 & 69 & ??? & - & - \\ 
%		& 2 & 127 & 142 & 95 & ??? & - & - \\ 
%		& 3 & 114 & 114 & 93 & ??? & - & - \\ 
%		\hline
%		quine-2 & 1 & 147 & 112 &  56 & ??? & - & - \\ 
%		& 2 & 161 & 123 & 101 & ??? & - & - \\ 
%		& 3 & 289 & 189 & 104 & ??? & - & - \\ 
%		\hline 
%'(I love you)-1 &  99 & 56 & 15 & 22 & ??? &  74 & 165 \\ 
%& 198 & 53 & 72 & 55 & ??? &  47 &  74 \\
%& 297 & 72 & 90 & 44 & ??? & 181 & 365 \\ 
%\hline
%'(I love you)-2 &  99 & 242 &  61 & 16 & ??? &  66 &  99 \\ 
%& 198 & 445 & 110 & 60 & ??? &  42 &  64 \\
%& 297 & 476 & 146 & 49 & ??? & 186 & 322 \\ 
%\hline 
	\end{tabular}
	\caption{The results of a quantitative evaluation: running times of 
	benchmarks in milliseconds}
	\label{compare-efficiency}
\end{table}

@@

In general, only \DFSi{} and \DFSbi{} consistently perform well. \DFSf{}
is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
implementations have obvious overhead in most cases. The reason might be that 
those strategies %tend to keep more intermediate answers in the memory.
tend to remember more states at the same time. 
Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is most 
resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
contrast, \DFSbi{} is still more stable than \DFSi{}. Thus, we consider 
\DFSbi{} a competitive alternative to \DFSi{}.
Among the two implementations of \BFS, \BFSimp{} constantly performs as well 
or better.

@@

We devise two new search strategies (i.e., balanced interleaving DFS 
(\DFSbi{}) and fair DFS (\DFSf{})) and devise a new 
implementation of \BFS{} (\BFSimp{}). These strategies have different features 
in fairness: \DFSbi{} has an almost-fair \disj{} and unfair \conj{}. 
\DFSf{} has fair \disj{} and unfair \conj{}. \BFS{} has both fair
\disj{} and fair \conj{}.

>>>>

%To summarize our contributions, we
%\begin{itemize}
%	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
%depth-first search (\DFSbi{}), a new search strategy with almost-fair \disj{}.
%	\item propose and implement \textbf{f}air depth-first search (\DFSf{}), 
%a new search strategy with fair \disj{}.
%	\item implement in a new way the standard breath-first search (\BFS) by 
%\citet{seres1999algebra}, a search strategy with fair \disj{} and fair 
%\conj{}. 
%We refer to \BFSser{} as their implementation and \BFSimp{} as our new 
%one. We formally prove that the two implementations are 
%semantically equivalent, however, \BFSimp{} runs faster in all benchmarks and 
%is shorter.
%\end{itemize}

Our research compares four search strategies with different features in fairness
(\autoref{table-fairness}). To summarize our contributions, we
\begin{itemize}
	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
	depth-first search (\DFSbi{})
	\item propose and implement \textbf{f}air depth-first search (\DFSf{})
	\item implement in a new way the standard breath-first search (\BFS) by 
	\citet{seres1999algebra}
	We refer to \BFSser{} as their implementation and \BFSimp{} as our new 
	one. We formally prove that the two implementations are 
	semantically equivalent, however, \BFSimp{} runs faster in all benchmarks 
	and 
	is shorter.
\end{itemize}

\begin{table}[h]
	\begin{tabular}{|c|c|c|}
		\hline 
		Search Strategies & disj & conj \\ 
		\hline 
		\DFSi & unfair & unfair \\ 
		\hline 
		\DFSbi & almost-fair & unfair \\ 
		\hline 
		\DFSf & fair & unfair \\ 
		\hline 
		\BFS & fair & fair \\ 
		\hline 
	\end{tabular} 
	\caption{Fairness of all search strategies}
	\label{table-fairness}
\end{table}

@@

% In this section, we define fair \disj{}, almost-fair \disj{} and fair 
%\conj{}. 
In this section, we define fairness levels in disjunction and conjunction
implementations. Specifically, a disjunction implementation can be
fair, almost-fair, or unfair. And a conjunction implementation can be
fair or unfair. Fairness, intuitively, measures how evenly a search strategy 
allocates computational resource to ``sibling'' spaces.

Before going further into fairness, we give a short review of the terms:
\emph{state}, \emph{space}, and \emph{goal}.
A \emph{state} is a collection of constraints. (Here, we restrict 
constraints to unification constraints.) Every answer corresponds to a 
state. A \emph{space} is a collection of states. And a \emph{goal} is a 
function from a state to a space.
Every state in the output space includes the input state and possibly more 
constraints.

@@

Although some strategies have a different level of fairness in \disj{}, they 
have similar behavior when there is no call to a relational definition 
in \conde{} clauses, including this case. (The only potential difference is 
that \DFSbi{} might reorder states.)

@@

Our solution requires a search strategy with \emph{fair} \conj{} to organize
states in buckets in spaces, where each bucket is a finite collection of states
and every space contains possibly infinite buckets,
and 
to allocate resources evenly among spaces derived from states in the 
same bucket. It is up to a search strategy designer to decide by what criteria 
to 
put states in the same bucket, and how to allocate resources among spaces 
related to different buckets.

@@

\begin{table}
	% run in command line, with a memory limit of 500MB
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline 
	benchmark & size & \DFSi & \DFSbi & \DFSf & \BFSimp{} & \BFSser{} \\
		\hline
		\veryrecursiveo & 100000 & 166 & 103 & 412 &   451 &   1024 \\
	    				& 200000 & 283 & 146 & 765 &   839 &   1875 \\
						& 300000 & 429 & 346 &2085 &  1809 &   4408 \\
		\hline  
		\appendo        & 	 100 &  17 &  18 &  17 &    17 &     65 \\
		         	  	& 	 200 & 145 & 137 & 137 &   142 &    121 \\
          		      	&	 300 & 388 & 384 & 387 &   429 &    371 \\
%   		& - & - & - & - & - & - \\
		\hline 
		\reverso		& 	  10 &   3 &   4 &   3 &    18 &     36 \\
						& 	  20 &  35 &  35 &  33 &  3070 &   3695 \\
						& 	  30 & 329 & 333 & 315 & 75079 & 107531 \\
		\hline
	\texttt{quine-1}	& 1 & 13 & 14 & 10 & - & - \\
						& 2 & 30 & 34 & 25 & - & - \\
						& 3 & 41 & 48 & 33 & - & - \\
		\hline
	\texttt{quine-2}	& 1 & 22 & 21 & 12 & - & - \\
						& 2 & 51 & 46 & 24 & - & - \\
						& 3 & 72 & 76 & 32 & - & - \\
		\hline
%		'(I love you)-1 &  99 &  10 &  37 & - &   40 &   96 \\
%						& 198 &  22 &  71 & - &  374 &  652 \\
%						& 297 &  24 & 292 & - & 1668 & 3759 \\
%		\hline
%		'(I love you)-2 &  99 & 485 & 179 & - &   44 &   94 \\
%						& 198 & 808 & 638 & - &  414 &  630 \\
%						& 297 &1265 & 916 & - & 1651 & 3691 \\ 
%		\hline 
\texttt{'(I love you)-1}&  999 &  78 &  70 &  59 & 254 & 605 \\
						& 1998 & 152 & 154 & 109 & 315 & 635 \\
						& 2997 & 430 & 440 & 266 & 317 & 643 \\
		\hline
\texttt{'(I love you)-2}&  999 & 631 & 173 &  61 & 253 & 605 \\
						& 1998 &1233 & 511 & 111 & 308 & 659 \\
						& 2997 &2145 & 682 & 269 & 307 & 642 \\ 
		\hline 
		% The data below is test with old benchmarks
%		quine-1 & 1 &  71 &  44 & 69 & ??? & - & - \\ 
%		& 2 & 127 & 142 & 95 & ??? & - & - \\ 
%		& 3 & 114 & 114 & 93 & ??? & - & - \\ 
%		\hline
%		quine-2 & 1 & 147 & 112 &  56 & ??? & - & - \\ 
%		& 2 & 161 & 123 & 101 & ??? & - & - \\ 
%		& 3 & 289 & 189 & 104 & ??? & - & - \\ 
%		\hline 
%'(I love you)-1 &  99 & 56 & 15 & 22 & ??? &  74 & 165 \\ 
%& 198 & 53 & 72 & 55 & ??? &  47 &  74 \\
%& 297 & 72 & 90 & 44 & ??? & 181 & 365 \\ 
%\hline
%'(I love you)-2 &  99 & 242 &  61 & 16 & ??? &  66 &  99 \\ 
%& 198 & 445 & 110 & 60 & ??? &  42 &  64 \\
%& 297 & 476 & 146 & 49 & ??? & 186 & 322 \\ 
%\hline 
	\end{tabular}
	\caption{The results of a quantitative evaluation: running times of 
	benchmarks in milliseconds}
	\label{compare-efficiency}
\end{table}

@@

%In general, only \DFSi{} and \DFSbi{} consistently perform well. \DFSf{}
%is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
%implementations have obvious overhead in most cases. The reason might be that 
%those strategies %tend to keep more intermediate answers in the memory.
%tend to remember more states at the same time. 
%Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is 
%most 
%resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
%\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
%contrast, \DFSbi{} is still more stable than \DFSi{}. Thus, we consider 
%\DFSbi{} a competitive alternative to \DFSi{}.
%Among the two implementations of \BFS, \BFSimp{} constantly performs as well 
%or better.

In general, variants of DFS usually performs better than BFS. The reason might 
be that \BFS{} tends to remember more states at the same time. Among the three 
variants of DFS, which all have unfair \conj{}, \DFSf{} is most resistant to 
clause permutation in \texttt{quine}s and \texttt{'(I love you)}s, followed by 
\DFSbi{} then \DFSi{}. Thus, we consider \DFSbi{} and \DFSf{} competitive 
alternatives to \DFSi{}. Among the two implementations of \BFS, \BFSimp{} 
constantly performs as well or better.

@@

We devise two new search strategies (i.e., balanced interleaving DFS 
(\DFSbi{}) and fair DFS (\DFSf{})) and devise a new 
implementation of \BFS{} (\BFSimp{}). These strategies have different features 
in fairness: \DFSbi{} has an almost-fair \disj{} and unfair \conj{}. 
\DFSf{} has fair \disj{} and unfair \conj{}. \BFS{} has both fair
\disj{} and fair \conj{}. No search strategy in this paper combines unfair 
\disj{} and fair \conj{}. This is because we haven't seen a case where this 
kind of search strategies would be interesting.
